### 2023-12-13 Tree.Rmd

---
title: "Tree-based"
author: "Mihir Kulkarni"
date: "2023-12-12"
output: html_document
---

### Explain tree model

### Pre-processing the data

```{r}
#Combine C and CL status into one variable and binarize
cirrhosis$Status <- ifelse(cirrhosis$Status == "C" | cirrhosis$Status == "CL", 1, 0)
```

# Split the data sets into training and testing

```{r}
#Set up the data

# Wrangle the Graduate data to set up training and testing datasets
modelCirrhosis <- cirrhosis %>%
  #drop_na() %>%
  mutate(
    tempID = row_number(),
    .before = Status
  )

## Set seed for reproducibility and slice ----
set.seed(380)
trainingData <- modelCirrhosis %>%
  group_by(Status) %>%
  slice_sample(prop = 0.8)

testingData <- modelCirrhosis %>%
  filter(!(tempID %in% trainingData$tempID))
```

### Grow tree via rpart

```{r}
# Grow Graduate tree via rpart package
rPartStatus <- rpart(
  formula = Status ~ N_Days + Drug + Age + Sex + Ascites + Hepatomegaly + Spiders + Edema + Bilirubin + Cholesterol + Albumin + Copper + Alk_Phos + SGOT + Platelets + Prothrombin + Stage,
  #Lost Triglycerides + 
  data = trainingData,
  method = "class",
  parms = list(split = "information")
  # We did not need to use the control parameters
)

```

### Plot tree

```{r}
# Display rpart.plot ----
# library(rpart.plot)
rpart.plot(
  x = rPartStatus,
  type = 2,
  extra = 101
)
```

### Plot tree

```{r}
# Using the rattle package to visualize the tree ----
fancyRpartPlot(
  model = rPartStatus,
  main = NULL,
  sub = NULL
)
```

### Plot cpTable

```{r}
invisible(capture.output({cpTable <- printcp(rPartStatus)}))

kable(
  x = cpTable,
  col.names = c("CP", "Num. of splits", "Rel. Error",
                "Mean Error", "Std. Deviation of Error"),
  digits = 3,
  booktabs = TRUE,
  align = "c",
  table.attr = 'data-quarto-disable-processing="true"'
)
```

### Plot cp visualization

```{r}
plotcp(
  x = rPartStatus,
  minline = TRUE,
  upper = "size"
)
```

### Prune tree

```{r}
# Prune the rpart Graduate Tree ----
rPartStatus2 <- prune(
  tree = rPartStatus,
  cp = 0.016
)
```

### Plot pruned tree

```{r}
fancyRpartPlot(
  model = rPartStatus2,
  main = NULL,
  sub = NULL
)
```

### Predictions

```{r}
pred_StatusRpart2 <- predict(
  object = rPartStatus2,
  newdata = testingData,
  type = "prob"
)

# Data Wrangling the predictions ----
StatusPrediction <- data.frame(
  rpart2_non_death = pred_StatusRpart2[, 1],
  rpart2_death = pred_StatusRpart2[, 2]
) %>%
  mutate(
    rpart2_pred = ifelse(
      test = rpart2_death > rpart2_non_death,
      yes = 1,
      no = 0
    )
  )

## Set predictions as factors
StatusPrediction$rpart2_pred <- as.factor(StatusPrediction$rpart2_pred)

# Merge supervision column into predictions data frame ----
StatusPrediction <- cbind(
  tempID = testingData$tempID,
  Status = testingData$Status,
  StatusPrediction
)
```

```{r}
StatusPrediction$Status <- factor(StatusPrediction$Status)

# Build confusion matrix for second tree model
conf_matrix <- conf_mat(
  data = StatusPrediction,
  truth = Status,
  estimate = rpart2_pred
)$table

kable(
  conf_matrix,
  col.names = c("Prediction/Supervision", "0", "1"),
  digits = 3,
  booktabs = TRUE,
  caption = "Model 1: Confusion Matrix",
  align = "c"
) %>%
kable_styling(latex_options = "HOLD_position")

accuracy <- accuracy(StatusPrediction, Status, rpart2_pred)
specificity <- specificity(StatusPrediction, Status, rpart2_pred)
sensitivity <- sensitivity(StatusPrediction, Status, rpart2_pred)
```
```{r}
# Build a data frame with model metrics ----
StatusPreds <- StatusPrediction %>%
  dplyr::select(tempID, Status, contains("_pred")) %>%
  pivot_longer(
    cols = !c(tempID, Status),
    names_to = "model",
    values_to = "prediction"
  )

accuracy <- StatusPreds %>%
  group_by(model) %>%
  accuracy(
    truth = Status,
    estimate = prediction
  )

sensitivity <- StatusPreds %>%
  group_by(model) %>%
  sensitivity(
    truth = Status,
    estimate = prediction,
    event_level = "second"
  )

specificity <- StatusPreds %>%
  group_by(model) %>%
  specificity(
    truth = Status,
    estimate = prediction,
    event_level = "second"
  )

modelMetrics <- bind_rows(
  accuracy,
  sensitivity,
  specificity
)
```
```{r}
# Make a nice looking table of model metrics ----
modelMetrics %>%
  dplyr::select(model, .metric, .estimate) %>%
  pivot_wider(
    id_cols = model,
    names_from = .metric,
    values_from = .estimate
  ) %>%
  kable(
    digits = 3,
    booktabs = TRUE,
    align = "c",
    table.attr = 'data-quarto-disable-processing="true"'
  )
```

```{r}
#Set up the data

# Wrangle the Graduate data to set up training and testing datasets
modelCirrhosis2 <- cirrhosis %>%
  drop_na() %>%
  mutate(
    tempID = row_number(),
    .before = Status
  )

## Set seed for reproducibility and slice ----
set.seed(380)
trainingData <- modelCirrhosis2 %>%
  group_by(Status) %>%
  slice_sample(prop = 0.8)

testingData <- modelCirrhosis2 %>%
  filter(!(tempID %in% trainingData$tempID))
```

# Comparison with logistic regression

```{r}
# Lower bound (Intercept only)
lower <- glm(
  formula = Status ~ 1,
  data = trainingData,
  family = binomial
)

# Upper bound 
upper <- glm(
  formula = Status ~ N_Days + Drug + Age + Sex + Ascites + Hepatomegaly + Spiders + Edema + Bilirubin + Cholesterol + Albumin + Copper + Alk_Phos + SGOT + Platelets + Prothrombin + Stage,
  data = trainingData,
  family = binomial
)

# Stepwise search
model2 <- step(
  object = lower,
  scope = list(
    lower = lower,
    upper = upper
  ),
  data = trainingData,
  direction = "both",
  k = 2
)
trainingResults <- trainingData
```

```{r}
# Store the predicted and actual values for Model 3:
trainingResults$model2Pred <- predict(model2, type = "response")
trainingResults$model2Class <- ifelse(trainingResults$model2Pred > 0.5, 1, 0)
trainingResults$Actual <- ifelse(trainingData$Status == 1, 1, 0)

# Create confusion matrix using table()
confusionMatrix <- table(Predicted = trainingResults$model2Class, Actual = trainingResults$Actual)

# Print and format confusion matrix
kable(confusionMatrix, booktabs = TRUE, align = "c")%>%kable_styling(latex_options = "HOLD_position")
```

```{r}
accuracy <- (confusionMatrix[1, 1] + confusionMatrix[0, 0]) / sum(confusionMatrix)
sensitivity <- confusionMatrix[1, 1] / (confusionMatrix[1, 1] + confusionMatrix[0, 1])
specificity <- confusionMatrix[0, 0] / (confusionMatrix[0, 0] + confusionMatrix[1, 0])

metrics_df <- data.frame(
  Metric = c("Accuracy", "Sensitivity", "Specificity"),
  Value = c(accuracy, sensitivity, specificity)
)

# Use kable to create the table
kable(metrics_df, format = "html", caption = "Model Performance Metrics")
```