### 2023-12-13 Regression.Rmd
yuh

---
title: "Regression-based"
author: "Mihir Kulkarni, Nithika Menon"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```