### 2023-12-14 Regression.Rmd
h

---
title: "Regression-based"
author: "Mihir Kulkarni, Nithika Menon"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Dataset for Regression
```{r}
cirrhosisRegression <- cirrhosis
```

```{r}

cirrhosisRegression <- cirrhosisRegression %>%
                                        filter(Status != "D") %>%
                                                          mutate(Transplant = ifelse(Status == 'C', 0, 1 ))

```

### Logistic regression model
```{r}

TransplantModel <- glm()

```