### 2023-12-14 EDA.Rmd
b

---
title: "EDA"
author: "Mihir Kulkarni, Nithika Menon"
date: "2023-12-12"
output: html_document
---

```{r}
#Required libraries

library(survival)
library(ggplot2)
library(survival)
library(survminer)

```

### Survival Analysis Based on Treatment and Stage: How does the survival rate differ among patients at different stages of PBC who received D-penicillamine versus placebo?

This question investigates the effectiveness of the drug D-penicillamine compared to a placebo, considering the stage of PBC. By analyzing survival rates across different disease stages and treatment types, we can assess the drug's effectiveness at various disease stages. This information is vital for clinicians to make informed decisions about treatment plans and for researchers to understand the drug's impact.

```{r}
surv_object <- Surv(cirrhosis$N_Days, cirrhosis$Status == 'D')
surv_fit <- survfit(surv_object ~ cirrhosis$Drug + cirrhosis$Stage, data = cirrhosis)
ggsurvplot(surv_fit, 
           data = cirrhosis,
           conf.int = TRUE, 
           #palette = c("#00AFBB", "#E7B800", "#FC4E07"),
           xlab = "Days", 
           ylab = "Survival Probability", 
           title = "Survival Curves by Treatment and Stage")
```

### Impact of Liver Complications on Survival: How do the presence of ascites, hepatomegaly, and spiders relate to survival time?

Exploring how liver-related symptoms (ascites, hepatomegaly, and spiders) affect patient survival provides insights into the severity of these complications in PBC progression. This can help in identifying high-risk patients and understanding the disease's impact on liver function.

```{r}
surv_fit_complications <- survfit(surv_object ~ cirrhosis$Ascites + cirrhosis$Hepatomegaly + cirrhosis$Spiders, data = cirrhosis)
ggsurvplot(surv_fit_complications, 
           data = cirrhosis,
           conf.int = TRUE, 
           #palette = c("#2E9FDF", "#FC4E07", "#6ACC65"),
           xlab = "Days", 
           ylab = "Survival Probability", 
           title = "Survival Curves by Liver Complications")
```

### Correlation between Biochemical Markers and Disease Stage: What is the correlation between biochemical markers (Bilirubin, Cholesterol, Albumin, Copper, Alk_Phos, SGOT, Triglycerides) and the histologic stage of the disease?

This analysis is crucial to understand how different biochemical markers correlate with the disease's progression. Such correlations can aid in the early detection of disease severity, help in monitoring the disease progression, and potentially guide treatment adjustments.

### Sex Distribution of Patients
```{r}
library(ggplot2)
library(dplyr)

cirrhosis$Sex <- factor(ifelse(cirrhosis$Sex == 0, "Female", "Male"))

# Preparing the data for pie chart
sex_data <- cirrhosis %>% 
            count(Sex) %>% 
            mutate(Percentage = n / sum(n) * 100)

# Pie Chart
ggplot(sex_data, aes(x="", y=Percentage, fill=Sex)) +
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
    scale_fill_brewer(palette="Pastel1") +
    theme_minimal() +
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank(),
          plot.title = element_text(size=14, face="bold")) +
    labs(fill="Sex", title="Distribution of Sex in PBC Patients")

```
Primary Biliary Cholangitis affects women much more than men at a 10-1 ratio. The data of this clinical study mimics the greater population. Investigating gender-based differences in disease progression can uncover any gender-specific patterns in PBC. This could lead to gender-tailored treatment approaches and a better understanding of the disease's biology, which might differ between males and females.

### Age Distribution of Patients

```{r}
# Converting age from days to years for better readability
cirrhosis$Age_Years <- cirrhosis$Age / 365.25

# Histogram
ggplot(cirrhosis, aes(x=Age_Years)) +
    geom_histogram(binwidth=5, fill="#69b3a2", color="#e9ecef") +
    theme_minimal() +
    labs(title="Age Distribution of PBC Patients", x="Age (years)", y="Count") +
    theme(plot.title = element_text(size=14, face="bold"))

```
The distribution of ages is quite normal.

### might have to mutate data below to filter out alive ppl but get the gist
```{r}

ggplot(data = cirrhosis, aes (Drug, Stage, fill = N_Days))  + geom_tile() + scale_fill_distiller(palette = "RdPu")

```

### Individuals with Edema and if they were on the drug
```{r}

library(ggmosaic)
cirrhosis %>%
  ggplot() +
  geom_mosaic(aes( x = product(Edema), fill = Drug))
#ggplot(data = cirrhosis) + geom_mosaic(aes(x = Drug , fill = Status))

```
Looking at the dist. of stages
```{r}

cirrhosis %>%
      ggplot(aes(x = Stage)) + geom_bar() # need color

```

```{r}
cirrhosis %>%
      ggplot(aes(x = Cholesterol, y = Bilirubin)) + geom_line()

```

### Comparison of Edema Status on Patient Outcomes

Edema is a common complication in liver diseases and can significantly impact patient quality of life and survival. Understanding how different levels of edema (none, controlled by diuretics, or persistent despite treatment) affect survival can inform patient management strategies and highlight the need for aggressive interventions in certain cases.