### 2023-12-14 Unsupervised.Rmd

---
title: "Unsupervised"
author: "Mihir Kulkarni"
date: "2023-12-12"
output: html_notebook
---

# Unsupevised Learning

 In this section, we sought to if there are sub-collections of patients. We expect sub-collections to mimic medical definitions of the stages of Biliary Cirrhosis but are curious to see if clustering can reveal some newer observations.

### Stages of Biliary Cirrhosis

Stage 1: There's inflammation and damange to the walls of medium-sized bile ducts.

Stage 2: There's blockage of the small bile ducts.

Stage 3: This stage marks the beginning of scarring.

Stage 4: Cirrhosis has developed. This permanent, severe, scarring and damange to the liver.

Source of information: https://www.healthline.com/health/primary-biliary-cirrhosis#stages

#do the ml, take about observational stages vs now quantify, look into the different variables and what they mean.

### Data Wrangling
```{r}

cirrhosisCluster <- cirrhosis

cirrhosisCluster <- cirrhosisCluster %>%                 
                                  select(-Stage)

cirrhosisCluster <- na.omit(cirrhosisCluster) #cannot have NA values in clustering

cirrhosisCluster$Age_Years <- round(cirrhosisCluster$Age_Years) #round age to whole numbers

#reode sex... recode others later if need be #female is 0

#cirrhosisCluster  <- cirrhosisCluster %>%
#  select( -c(Status, Drug, Edema)) %>%

# mutate(Sex = ifelse(Sex == 'Female', 0, 1))

  cirrhosisCluster  <- cirrhosisCluster %>%
    mutate(Sex = ifelse(Sex == 'Female', 0, 1)) %>%
          mutate(Transplant = ifelse(Status == "CL", 1, 0)) %>%
              mutate(Status = ifelse(Status %in% c('C', 'CL'), 0 , 1)) %>%
                    mutate(Drug = ifelse(Drug == "D-penicillamine", 0, 1)) %>%
                        mutate(EdemaDiurectics = ifelse(Edema %in% c('S', 'Y'), 1, 0)) %>%
                              mutate(NoEdemaORD = ifelse(Edema == 'N' , 1, 0)) %>%
                                        mutate(EdemaANDD = ifelse(Edema == "Y", 1, 0)) %>%
                                                mutate(EdemaORD = ifelse(Edema == "S", 1, 0)) %>%
                                                                              select(-Edema)

#Data is already factored

#mutate(Sex = ifelse(Sex == 'Female', 0, 1)) %>%

   #mutate(Status = ifelse(Status %in% c('C', 'CL'), 0 , 1)) %>%
          #    mutate(Drug = ifelse(Drug == "D-penicillamine", 0, 1))
```

#make column yes no edema
#then yes no under treatment

#same for tranplant stuff

```{r}
distCirrhosis <- dist(
  x = cirrhosisCluster,
  method = "euclidean"
)

```

### A mixed hierarchial and non-hierarchial

```{r}

hybridCirrhosis <- hkmeans(
  x = cirrhosisCluster,
  k = 4,
  hc.metric = "euclidean",
  hc.method = "ward.D",
  iter.max = 10
)

```

### Visualize Results

### Making a color palette
```{r}

StagesPalette <- c("#AA336A", "#770737", "#40B5AD", "#009E60")

```

```{r}

## MAKE A NEW PALLETE TO VISUALIZE

# Plot the initial dendrogram for hybrid approach ----
set.seed(380)
hkmeans_tree(
  hkmeans = hybridCirrhosis,
  rect.col = StagesPalette,
  cex = 0.4,
  main = "Initial Hierarchical Clusters"
)

```

```{r}
# Plot the final dendrogram for hybrid approach ----
# library(factoextra)
fviz_dend(
  x = hybridCirrhosis,
  cex = 0.4,
  palette = StagesPalette,
  rect = FALSE,
  horiz = TRUE,
  repel = TRUE,
  main = "Final Dendrogram"
)

```

```{r}

# Plot the final clustering for hybrid approach ----
# library(factoextra)
fviz_cluster(
  object = hybridCirrhosis,
  stand = FALSE,
  geom = "point",
  main = "Hybrid Cluster Plot"
) +
  scale_color_manual(values = StagesPalette) +
  scale_fill_manual(values = StagesPalette) +
  theme_bw()

```