### 2023-12-15 Regression.Rmd
Re-did logistic regression with new goals

---
title: "Regression-based"
author: "Mihir Kulkarni, Nithika Menon"
date: "2023-12-12"
output: html_document
---

# Part II: Exploring the predictors for a Liver Transplant using logistic regression

Next, let us explore our question with a different type of model. To do this, we will be using (binary) logistic regression model

## Introduction

Logistic regression is a statistical method used for binary classification, which predicts the probability of an outcome that can be either true or false. This is done by understanding the relationship between a dependent binary variable and one or more independent variables. Logistic regression is easy to implement and interpret. However there are also some drawbacks, the model assumes a linear relationship between the independent variables and the log odds of the dependent variable, which may not always hold true in complex real-world scenarios. Furthermore, unlike the decision tree, linear regression models cannot ignore N/A values.

![Logistic Regression Model ([Source](https://www.spiceworks.com/tech/artificial-intelligence/articles/what-is-logistic-regression/))](46-4.png)

## Methodology

Similar to the tree, we will start by splitting the data set into training and testing sets. Note that the data now only contains instances where the patient was deceased. The training set will be used to train our model, while the testing set will help evaluate its performance. We'll use 80% of the data for training and the remaining 20% for testing. To allow reproducible code, we have fixed the seed at 380.

```{r}
cirrhosisRegression <- cirrhosis
cirrhosisRegression$Status <- ifelse(cirrhosisRegression$Status == "C" | cirrhosisRegression$Status == "CL", 1, 0)
```

### Logistic regression model

```{r}
#model data
LRmodelData <- cirrhosisRegression %>%
  drop_na() %>%
  mutate(
    tempID = row_number(),
    .before = Status
  )

# Set seed for reproducibility and slice
set.seed(380)
trainingData <- LRmodelData %>%
  group_by(Status) %>%  # group_by() function ensures that the data
  slice_sample(prop = 0.75)

testingData <- LRmodelData %>%
  filter(!(tempID %in% trainingData$tempID))

trainingResults <- trainingData
```

```{r}
# Form Candidate Model 1
model1 <- glm(
  formula = Status ~ Platelets,
  data = trainingData,
  family = binomial
)
```

```{r models}
# Lower bound (Intercept only)
lower <- glm(
  formula = Status ~ 1,
  data = trainingData,
  family = binomial
)

# Upper bound 
upper <- glm(
  formula = Status ~ Drug + Age + Sex + Ascites + Hepatomegaly + Spiders + Edema + Bilirubin + Cholesterol + Albumin + Copper + Alk_Phos + SGOT + Platelets + Prothrombin + Stage,
  data = trainingData,
  family = binomial
)

# Stepwise search
model2 <- step(
  object = lower,
  scope = list(
    lower = lower,
    upper = upper
  ),
  data = trainingData,
  direction = "both",
  k = 2
)
```

# Results

Initially, we'll delve into the two preliminary models independently to
understand where they stand. Following that, we'll be deploying the best
candidate model on our test data. Regarding confusion matrices, we'll
employ a basic rule: if the predicted probability of a student
graduating exceeds 0.5, we'll categorize them as having graduated (naïve
rule).

## Model 1

```{r}
# Model 1 Coefficient Table
as.data.frame(summary(model1)$coefficients) %>%
  rownames_to_column(var = "term") %>%
  rename(coefficient = Estimate) %>% 
  mutate(
    prob_odds = case_when(
      coefficient == "(Intercept)" ~ exp(coefficient)/(1 + exp(coefficient)),
      .default = exp(coefficient)
    ),
    .after = coefficient
  ) %>%
  mutate(
    `Pr(>|z|)` = ifelse(
      test = `Pr(>|z|)` < 0.001,
      yes = paste("< 0.001"),
      no = `Pr(>|z|)`
    ),
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      grepl(x = term, pattern = "Platelets") ~ "Platelets"
    )
  ) %>%
  kable()

```

```{r}
library(janitor)
# Building confidence intervals for Model 1 coefficients
model1CI <- confint(
  object = model1,
  parm = "Platelets",
  level = 0.9
)

trainingResults <- trainingData %>%
  ungroup() %>%
  mutate(model1Pred = predict(object = model1, newdata = ., type = "response"))

# Apply naïve rule ----
trainingResults <- trainingResults %>%
  mutate(
    model1Class = case_when(
      model1Pred > 0.5 ~ 1,
      .default = 0
    )
  )

#Confusion Matrix for Model 1
trainingResults %>%
  mutate(Patient_status = ifelse(Status == 1, 1, 0)) %>%
  tabyl(var1 = model1Class, var2 = Patient_status) %>%
  adorn_title(
    placement = "combined",
    row_name = "Predicted",
    col_name = "Actual"
  ) %>%
  kable(
    booktabs = TRUE,
    align = "c",
    caption = "Model 1 Confusion Matrix"
  )

```

Next let us look at our Model 2, which has multiple factors as discussed
earlier.

```{r}
#Coeff for model 2
as.data.frame(summary(model2)$coefficients) %>%
  rename(coefficient = Estimate) %>% 
  mutate(
    prob_odds = case_when(
      coefficient == "(Intercept)" ~ exp(coefficient)/(1 + exp(coefficient)),
      TRUE ~ exp(coefficient)
    ),
    .after = coefficient
  ) %>%
  kable()

```

```{r}
#do the Tukey-Anscombe plot
ggplot(
  data = data.frame(
    residuals = residuals(model2, type = "pearson"),
    fitted = fitted(model2)
  ),
  mapping = aes(x = fitted, y = residuals)
) +
  geom_point() +
  geom_smooth(
    formula = y ~ x,
    method = stats::loess,
    method.args = list(degree = 1),
    se = FALSE,
    linewidth = 0.5
  ) +
  theme_bw() +
  labs(
    x = "Fitted",
    y = "Pearson Residuals"
  )
```

```{r}
#plot the gvif
as.data.frame(car::vif(model2)) %>%
  kable(
    digits = 3,
    align = "lcccc",
    booktab = TRUE,
    format.args = list(big.mark = ","),
    table.attr = 'data-quarto-disable-processing="true"',
    label = "GVIF analsyis"
  )

```

Here we can see the generalized variance inflation factors for candidate
Model 2. Multicollinearity is a statistical concept where several
independent variables in a model are correlated and this is an important
consideration especially when we deal with so many categories.

```{r}
#Store the predicted and actual values for Model 2
trainingResults$model2Pred <- predict(model2, type = "response")
trainingResults$model2Class <- ifelse(trainingResults$model2Pred > 0.5, 1, 0)
trainingResults$Actual <- ifelse(trainingData$Status == 1, 1, 0)

# Create confusion matrix using table
confusionMatrix <- table(Predicted = trainingResults$model2Class, Actual = trainingResults$Actual)

kable(confusionMatrix, caption = "Confusion matrix for Model 2") %>%
  kable_classic(latex_options = "HOLD_position")

```

Lastly, let us look at the separation plots for each of the models.

```{r}
library(pROC)
library(separationplot)
# Fit ROC Curves for later
## Model 1
model1ROC <- roc(
  formula = Status ~ model1Pred,
  data = trainingResults
)
model1ROC_df <- data.frame(
  threshold = model1ROC$thresholds,
  sensitivity = model1ROC$sensitivities,
  specificity = model1ROC$specificities,
  model = "Model 1"
)
## Model 2
model2ROC <- roc(
  formula = Status ~ model2Pred,
  data = trainingResults
)
model2ROC_df <- data.frame(
  threshold = model2ROC$thresholds,
  sensitivity = model2ROC$sensitivities,
  specificity = model2ROC$specificities,
  model = "Model 2"
)
```

```{r}
# Convert 'Actual' column to numeric 0/1
trainingResults <- trainingResults %>%
  mutate(
    actualNum = if_else(Actual == 0, 0, 1)
  )

#Sepeation Plot
par(mar = c(4,0,0,0))
separationplot(
  pred = trainingResults$model1Pred, 
  actual = trainingResults$actualNum, 
  type = "rect",
  line = TRUE, 
  lwd2 = 2,
  show.expected = TRUE, 
  newplot = FALSE,
  heading = "Model 1"
)

```

```{r}
#Sepeation Plot
par(mar = c(4,0,0,0))
separationplot(
  pred = trainingResults$model2Pred, 
  actual = trainingResults$actualNum, 
  type = "rect",
  line = TRUE, 
  lwd2 = 2,
  show.expected = TRUE, 
  newplot = FALSE,
  heading = "Model 2"
)
```

The separation plot assesses the the fit of the model by providing the
model's ability to predict occurrences with a high probability and
non-occurrences with low probability. The separation plot above
separation plot suggests that Model 2 has a reasonably good performance
in predicting graduates, especially for the observations on the
left-most side of the plot compared to Model 1 with the training data.
We will later compare this graph with the testing data.

Lastly, let us look at the ROC curves for bothe the models.

```{r}
## Merge into existing data frame
rocData <- rbind(model1ROC_df, model2ROC_df)

## AUC Data
aucData <- data.frame(
  model = c("Model 1", "Model 2"),
  auc = c(model1ROC$auc, model2ROC$auc)
)
```

```{r}
#ROC plot
ggplot(
  data = rocData,
  mapping = aes(x = 1 - specificity, y = sensitivity, color = model)
) +
  geom_path() +
  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dotted"
  ) +
  geom_text(
  inherit.aes = FALSE,
  data = aucData,
  mapping = aes(label = paste(model, "AUC: \n", round(auc, 3))),
  x = c(0.25, 0.15),
  y = c(0.4, 1.05)
)
```

From the graphs we can interpret that:

Model 1: - Its ROC curve is above the line of no discrimination,
indicating that the model has some predictive capabilities. - The AUC is
0.621, which is better than random guessing but suggests there's room
for improvement.

Model 2: - The ROC curve for Model 2 is significantly above that of
Model 1, and much closer to the top-left corner, indicating better
predictive performance. - The AUC is 0.768, which suggests a "good"
classification performance, and it's notably better than Model 1. - Its
ability to discriminate between positive and negative classes is
superior to that of Model 1.

Lastly, let us plot our influence plot.

```{r}
# Influence Plot for Model 2
idCases <- car::influencePlot(model2)
```

The influence plot shows several data points with high leverage and
large residuals, indicating potential outliers. Some observations,
notably those labeled like "23881," have significant influence on the
regression model due to their Cook's D values.

# Testing our model

Using our final model, we now turn to see how well this classifier does
on our testing data. Recall that we initally set the test data during
the train test split.

The confusion matrix below shows the performance of our model using the
naïve decision rule.

```{r}
# Set up testing data results
testingData <- testingData %>%
  mutate(
    gradNum = case_when(
      Status == 0 ~ 0,
      Status == 1 ~ 1
    ),
    .after = Status
  )
testingData$predict <- predict(
  object = model2,
  newdata = testingData,
  type = "response"
)
testingData <- testingData %>%
  mutate(
    model2Class = case_when(
      predict > 0.5 ~ 1,
      .default = 0
    )
  )
```

```{r}
testingData$Status <- ifelse(testingData$Status == 1, 1, 0)

# Build Confusion Matrix for Testing Data
testingData %>%
  tabyl(var1 = model2Class, var2 = Status) %>%
  adorn_title(
    placement = "combined",
    row_name = "Predicted",
    col_name = "Actual"
  ) %>%
  kable(
    caption = "Confusion Matrix for Test data"
  )
```

From the matrix, we can see that the model has the following metrics:

-   True Positive (TP): Predicted Graduated and actually Graduated. In
    this case, 7251.
-   False Positive (FP): Predicted Graduated but actually Not Graduated.
    In this case, 3505.
-   False Negative (FN): Predicted Not Graduated but actually Graduated.
    In this case, 1936.
-   True Negative (TN): Predicted Not Graduated and actually Not
    Graduated. In this case, 5395.

From this, we can interpret that.

-   Accuracy: Approximately 69.92%
-   Precision: Approximately 67.41%
-   Recall (Sensitivity): Approximately 78.93%
-   F1 Score: Approximately 72.72%

This shows that our model is still capable with new data, with a slight
reduction to the overall performance.

```{r}
#Sepeation Plot
par(mar = c(4,0,0,0))
separationplot(
  pred = testingData$predict, 
  actual = testingData$gradNum, 
  type = "rect",
  line = TRUE, 
  lwd2 = 2,
  show.expected = TRUE, 
  newplot = FALSE
)

```