### 2023-12-15 Submission.Rmd

---  
title: "Submission"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---

```{r child = 'Final Project.Rmd'}
```

```{r child = 'EDA.Rmd'}
```

```{r child = 'Tree.Rmd'}
```

```{r child = 'Regression.Rmd'}
```

```{r child = 'Unsupervised.Rmd'}
```