### 2023-12-15 Final Project.Rmd
Wrote the introduction for the data

---
title: "Final Project Report: An exploration of Biliary Cirrhosis and Treatments"
output: html_notebook
authors: "Mihir Kulkarni, Nithika Menon"
---

# Background

## What is Biliary Cholangitis?

Primary Biliary Cholangitis (previously called Primary Biliary Cirrhosis) is an autoimmune disease where bile ducts become swollen and inflamed and block the flow of bile. Bile is a substance that aids with digestion. The bile ducts carry bile from the liver to the small intestine. The swelling and inflammation can lead to scarring of the liver which is cirrhosis. Advanced cirrhosis can lead to liver failure or liver cancer. Medication can slow progression. There is no definite cure at this time.

![[Primary Biliary Cholangitis ([Source](https://www.google.com/search?client=safari&sca_esv=591276386&rls=en&sxsrf=AM9HkKmtJVOC2Jrf2BSYVB2O-QcnTSgx6g:1702664705100&q=Primary+Biliary+Cholangitis&tbm=isch&source=lnms&sa=X&ved=2ahUKEwid0qyniJKDAxX9FFkFHVH3CcoQ0pQJegQIDxAB&biw=1440&bih=848&dpr=2#imgrc=UeV_fhP-8vPOlM))]{.underline}](17715-primary-biliary-cholangitis.webp)

## Why is the data important to us?

Both of us have been very interetsted in the intersection of medicine and machine learning. The idea that the concepts we are learning in our DS majors could have such an impact on the health industry was really special. Another important aspect was that there is so much potential for advancement of medicine, we both agreed that this area is MLs most promising aspect. The cirrhosis data set perticularly stood out to us because the data set included so many interesting attributes and we anted to see how they are connected with the outcomes of the patient.

The data set also includes some important limitations. Most importantly, there are only 418 cases, greatly affecting the types of analysis we could do. This shows realm world problems data scientists face. Thus, our goal is to use our Statistical DS skillset to approach this problem, despite the limitations.

## Objectives

Exploring the study, we concluded on two core questions we hope to address in this report:

### CRQ1: What predictors, if any affected the status of the patient at the end of the study?

### CRQ2: Are there factors not currently included in medical definition that can help distinguish different stages of Biliary Cholangitis?

## Clinical Study + Dataset Background :

The dataset is produced from a clinical study by Mayo Clinic run from 1974-1984. The final 'Status' of each patient was observed in 1986. 'Status' was either Dead, Censored, or Censored due to liver transplant. A link to the original study is here: <https://faculty.washington.edu/abansal/ShortCourse_DynamicDecisionMaking/Dickson1989_MayoPBCOriginalArticle.pdf>

The data contains the following attributes ([Data Source](https://www.kaggle.com/datasets/fedesoriano/cirrhosis-prediction-dataset/data)):

1) **ID**: unique identifier\
2) **N_Days**: number of days between registration and the earlier of death, transplantation, or study analysis time in July 1986\
3) **Status**: status of the patient C (censored), CL (censored due to liver tx), or D (death)\
4) **Drug**: type of drug D-penicillamine or placebo\
5) **Age**: age in [days]\
6) **Sex**: M (male) or F (female)\
7) **Ascites**: presence of ascites N (No) or Y (Yes)\
8) **Hepatomegaly**: presence of hepatomegaly N (No) or Y (Yes)\
9) **Spiders**: presence of spiders N (No) or Y (Yes)\
10) **Edema**: presence of edema N (no edema and no diuretic therapy for edema), S (edema present without diuretics, or edema resolved by diuretics), or Y (edema despite diuretic therapy)\
11) **Bilirubin**: serum bilirubin in [mg/dl]\
12) **Cholesterol**: serum cholesterol in [mg/dl]\
13) **Albumin**: albumin in [gm/dl]\
14) **Copper**: urine copper in [ug/day]\
15) **Alk_Phos**: alkaline phosphatase in [U/liter]\
16) **SGOT**: SGOT in [U/ml]\
17) **Triglycerides**: triglicerides in [mg/dl]\
18) **Platelets**: platelets per cubic [ml/1000]\
19) **Prothrombin**: prothrombin time in seconds [s]\
20) **Stage**: histologic stage of disease (1, 2, 3, or 4)

For the basic pre-processing, we binary-encoded the categorical variables in the data set and set them as factors.

```{r load packages}

library(survival)
library(ggplot2)
library(survival)
library(survminer)
library(kableExtra)
library(factoextra)

```

```{r}
#load in the data
library(tidyverse)
library(kable)
#load data
cirrhosis <- read_csv("cirrhosis.csv")
```

```{r}
#encode the sex, ascites, hepatomegaly, spiders, edema into binary variables
cirrhosis$Sex <- ifelse(cirrhosis$Sex == "F", 0, 1)
cirrhosis$Ascites <- ifelse(cirrhosis$Ascites == "Y", 1, 0)
cirrhosis$Hepatomegaly <- ifelse(cirrhosis$Hepatomegaly == "Y", 1, 0)
cirrhosis$Spiders <- ifelse(cirrhosis$Spiders == "Y", 1, 0)
#factorize the stage
cirrhosis$Stage <- factor(cirrhosis$Stage)
cirrhosis$Status <- factor(cirrhosis$Status)
cirrhosis$Drug <- factor(cirrhosis$Drug)
cirrhosis$Sex <- factor(cirrhosis$Sex)
#cirrhosis$Ascites <- factor(cirrhosis$Ascites)
#cirrhosis$Hepatomegaly <- factor(cirrhosis$Hepatomegaly)
#cirrhosis$Spiders <- factor(cirrhosis$Spiders)
cirrhosis$Edema <- factor(cirrhosis$Edema)
cirrhosis$Sex <- factor(ifelse(cirrhosis$Sex == 0, "Female", "Male"))

#Drop the first column
cirrhosis <- cirrhosis[,-1]
```

-   Engage in Exploratory Data Analysis

-   Use Machine Learning Algorithms:

    -   You must use at least one of each: a Regression-based algorithm, a Tree-based algorithm, and an Unsupervised Learning algorithm.
    -   Each algorithm must be explained for a general audience.
    -   All results should be explained.
    -   Where appropriate, comparisons between methods should be discussed. This includes highlighting the strengths and weaknesses of the various approaches as well as when each is appropriate.

-   A discussion of what you've learned from your work about the context and recommendations for future work.

-   Author Contributions--list out who did what in the project.

-   References--Citation style is your choice, but all sources should be documented (both in text and in the References section). This includes where you got your data.

-   A commented Code Appendix

```{r}

cirrhosis %>% 
  ggplot(aes(x = N_Days)) + geom_histogram() +
  facet_wrap(facets = vars(Status))

```